---
title: ""
author: Emil Frej Brunbjerg
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
  github_document:
    toc: true
bibliography: references.bib
---

# What is this document?

@gallagher developed two metrics to identify who would respond best to an
experimental type of cancer therapy. These two metrics determine what
cancer patients would benefit more from receiving *adaptive therapy*
rather than standard care. They explicitly state that they leave the
quantification of uncertainty for these metrics for future work. In this
document, **I demonstrate how to quantify the uncertainty for one of these
measures: the *Delta AT Score*.** The Delta AT Score is the estimated relative benefit in time-to-progression of choosing adaptive therapy
over standard treatment for an individual patient. Using clinical data of a single patient, I demonstrate how one could go about estimating the uncertainty of this measure. This document should therefore
be considered a case-study, but it demonstrates the feasibility of the
approach. Applying it to multiple patients and other similar measures
should be straightforward.

Additionally, I investigate how the uncertainty in estimating the Delta AT Score
changes for each observation made for the specific patient. The trajectory
in uncertainty makes me somewhat skeptical that the uncertainty in delta
AT can reliably be decreased to clinically useful levels simply by adding
more observations. Instead, I think it might be necessary to adopt ideas
from sequential decision-making under uncertainty such as balancing
exploitation and exploration to adequately deal with the uncertainty in
estimates. This could be done by partly optimizing for observations that
reduce uncertainty, such that observations that provide large reductions in uncertainty are actively sought.

# What is Adaptive Therapy?

Adaptive therapy attempts to stabilize cancer as a chronic disease rather than
eradicate it. This is done by considering the evolutionary forces that
treatment places on cancer cells [@Adaptive].
To my understanding, the idea is that a tumour is typically limited by a carrying
capacity, and its cells therefore compete against each other. Cells that proliferate faster will tend to
crowd out their more slowly dividing cousins. Assuming that there is some fitness
cost to developing resistance against treatment, treatment-sensitive
cells should crowd out treatment-resistant cells.

It is only when treatment is applied that resistant cells have an
advantage, and if treatment is applied for long enough, the resistant
cells will take over control of the tumor. Adaptive therapy
attempts to strike a balance between killing off sensitive cells to keep
the cancer in check, but not killing so many that the resistant cells
will take over.

It is my understanding that the adaptive therapy research program has
mainly focused on prostate cancer. I have therefore done the same in this
document.

## Prostate Cancer as a Lotka-Volterra Model

Lotka-Volterra models are simple coupled ordinary differential equations
that describe how multiple species grow given the number of individuals
at each time point.

I don't have a complete overview, but I've found that at least two
different Lotka-Volterra models are used throughout adaptive therapy
literature to model treatment trajectories for prostate cancer. There
might be more. But they share the hurdle that they only produce a signal of the total cell population. In other words, one cannot directly
observe the number of sensitive and resistant cells. Only the biomarker
Prostate Specific Antigen (PSA) is measured, and is thought to be a
signal based on the total number of cancer cells.

## Observations from a Cognitive Science Student

My background is in cognitive science and, beyond likely
misunderstanding crucial aspects of oncology, it has made me note some
parallels between mathematical modelling in adaptive therapy and models
of cognition. Given a forward model, e.g. a Lotka-Volterra
model, one should be able to infer the underlying parameters of the model
and thereby predict individual outcomes. I'm mostly familiar with
solving this problem using Bayesian inference. Luckily, a Finnish
master's thesis @forslund showed that this is feasible for at least one
specific formulation of the Lotka-Volterra model for prostate cancer.

Bayesian inference produces distributions over plausible parameter
values instead of point estimates typical of maximum likelihood procedures. The estimate is a distribution, constructed by combining the likelihood function with prior distributions placed on the free parameters.

Assuming that uncertainty in estimates is substantial, in my eyes planning adaptive
therapy for individual patients is a type of sequential decision-making under
uncertainty. Explore-exploit dynamics could apply, and planning treatment
might benefit from balancing reward maximization and information gain. In this light, deciding when to
treat would also depend on the potential information gain from applying
treatment.

# Quantifying Uncertainty

## Patient 12

The following data are PSA (biomarker of total cell count) readings normalized by the first
observation from a prostate cancer patient undergoing an *intermittent* therapy trial by
@bruchovsky2006. I'm in doubt whether practitioners would classify this as adaptive therapy, but it is a data set that is frequently used to model adaptive therapy. For example @gallagher, who developed the Delta AT Score that I'm estimating the uncertainty for, use the same data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 11, fig.height = 5)
library(patchwork)
library(kldest)
source("kl_divergence_p12_fitting.R")
```

```{r psa_plot, fig.cap="*Treatment trajectory for patient 12. Dots indicate Prostate Specific Antigen measurements and shaded regions indicate that treatment was given.*"}
p_psa <- ggplot() +
  geom_rect(data = trt_abs, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "steelblue", alpha = 0.12, inherit.aes = FALSE) +
  geom_point(data = psa_data, aes(x = day, y = psa_norm),
             colour = "black", size = 2) +
  scale_colour_manual(values = c("S" = "steelblue", "R" = "tomato", "S+R" = "grey30"),
                      name = NULL) +
  labs(x = "Day", y = "Normalized PSA") +
  theme_minimal()

p_psa
```

## The Mathematical Model

As mentioned above, one way to quantify this uncertainty is to use the
forward Lotka-Volterra model to conduct Bayesian inference to find
distributions over parameters given data. I will do exactly that for
Patient 12. They derive their measures from the following Lotka-Volterra
model (equation 1 in their paper)

$$
\begin{align*}
\frac{dS}{dt} &= r_SS (1 - \frac{S+R}{K})(1-d_DD)-dS \\
\frac{dR}{dt} &= r_RR(1 - \frac{S+R}{K}) - dR\end{align*}
$$

$dS$ is the change in sensitive cells $S$ at time $t$, and $dR$ is the
change in resistant cells $R$. $r_S$ and $r_R$ are proliferation rates
of sensitive and resistant cells. $K$ is the carrying capacity of the
tumour, so the growth of the cells depends on how much of the shared
carrying capacity is taken up. $d_S$ and $d_R$ are the death rates of
sensitive and resistant cells. $d_D$ indicates the extra death rate of
the sensitive cells when the treatment indicator $D$ is on.

@gallagher then reparameterize the model and fix values of certain
parameters ($K$, $d_D$ and $r_S$) due to some tractability concerns that
I won't pretend to understand. The takeaway is that the final model only
has the following four free parameters:

-   $cost$*.* The penalty in proliferation rate that resistant cells
    suffer compared to sensitive cells.

-   $turnover$*.* The rate at which both resistant and sensitive cells
    die off.

-   $N_0$. The normalized amount of total cancer cells at $t_0$.

-   $R_0$. The fraction of resistant cells at $t_0$

Given the probabilistic modelling approach, I also have to specify from
what distribution the observations are sampled. For simplicity, I've
assumed that normalized Prostate Specific Antigen observations show
Gaussian noise around the normalized total cell count

$$PSA_t \sim N(S_t+R_t, \sigma^2)$$

which leaves us with another free parameter:

-   $\sigma^2$. The amount of noise in the Prostate Specific Antigen
    measurements.

By writing up the reparameterized Lotka-Volterra equation in the
probabilistic programming language *Stan* I can sample parameter values
according to how probable they are given a prior, the data and the
chosen likelihood function. While the fitted model generally follows the
trend in the data points, many observations fall outside the 95%
credible interval. This is a red flag for the adequacy of the
mathematical model and priors, since it is unlikely to produce
observations similar to the actual data. However, it is the same mathematical model that
@gallagher use.

```{r fitted_cells, fig.height=5, fig.cap="*Treatment trajectory for patient 12. Dots indicate PSA measurements and shaded regions indicate that treatment was given. Black line is the median value for $S+R$ at each time point. Grey ribbons represent 95% credible intervals.*"}
max_fit <- readRDS(file.path(CACHE_DIR, "p12_t28.rds"))

cell_draws <- spread_draws(max_fit, S_pred[i], R_pred[i]) %>%
  mutate(day   = psa_data$day[i],
         total = S_pred + R_pred)

cell_sum <- cell_draws %>%
  group_by(i, day) %>%
  summarise(
    S_med = median(S_pred), S_lo = quantile(S_pred, 0.025), S_hi = quantile(S_pred, 0.975),
    R_med = median(R_pred), R_lo = quantile(R_pred, 0.025), R_hi = quantile(R_pred, 0.975),
    T_med = median(total),  T_lo = quantile(total,  0.025), T_hi = quantile(total,  0.975),
    .groups = "drop"
  )

p_pred <- ggplot(cell_sum, aes(x = day)) +
  geom_rect(data = trt_abs, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "steelblue", alpha = 0.12, inherit.aes = FALSE) +

  geom_ribbon(aes(ymin = T_lo, ymax = T_hi), fill = "grey40",    alpha = 0.15) +
  geom_line(aes(y = T_med, colour = "S+R"), linewidth = 0.9) +
  geom_point(data = psa_data, aes(x = day, y = psa_norm),
             colour = "black", size = 2) +
  scale_colour_manual(values = c("S" = "steelblue", "R" = "tomato", "S+R" = "grey30"),
                      name = NULL) +
  labs(x = "Day", y = "Normalized cell count / PSA") +
  theme_bw()


p_pred
```

## Inferring Parameter Estimates and Uncertainties

Fitting the Lotka-Volterra system used by @gallagher to patient 12's
normalized Prostate Specific Antigen observations yields a probability
distribution over the parameters of the reparameterized Lotka-Volterra
system. For example, there is substantial uncertainty in estimating the cost parameter.

```{r param_posteriors, fig.height=6}
param_draws <- spread_draws(readRDS(file.path(CACHE_DIR, "p12_t28.rds")),
                            cost, turnover, n0, rFrac, sigma) %>%
  select(cost, turnover, n0, rFrac, sigma) %>%
  mutate(delta_at = calc_delta_at(turnover * rS, k, K,
                                     (1 - cost) * rS, n0 * rFrac)) %>%  #dR, k, K, rR, R0

  pivot_longer(everything(), names_to = "param", values_to = "value")

prior_param <- tibble(
  cost     = rbeta(N_PRIOR, 1, 1),
  turnover = rbeta(N_PRIOR, 2, 5),
  n0       = rbeta(N_PRIOR, 1, 1),
  rFrac    = rbeta(N_PRIOR, 1, 30),
  sigma    = rlnorm(N_PRIOR, -2, 1)
) %>%
  pivot_longer(everything(), names_to = "param", values_to = "value")

param_draws %>%
  filter(param != "delta_at") %>%
ggplot(aes(x = value)) +
  geom_density(data = prior_param, aes(x = value, colour = "Prior", fill = "Prior"),
               linetype = "dashed", linewidth = 0.7, alpha = 0.0) +
  geom_density(aes(colour = "Posterior", fill = "Posterior"), alpha = 0.4, linewidth = 0.8) +
  facet_wrap(~param, scales = "free", ncol = 3) +
  scale_colour_manual(name = NULL, values = c("Prior" = "black", "Posterior" = "steelblue")) +
  scale_fill_manual(name = NULL,   values = c("Prior" = "black", "Posterior" = "steelblue")) +
  labs(x = NULL, y = "density") +
  theme_classic() +
  theme(legend.position  = c(.87, .18),
        legend.justification = c(.9, 0))
```

By using the posterior samples, I can propagate the uncertainty in
the parameter estimates through to the Delta AT Score. The Delta AT Score depends on the Lotka-Volterra parameters (see the original paper for the mathematical
details). Since Stan provides thousands of parameter samples, I calculate the Delta AT Score for each draw and use the resulting distribution as my estimate.

```{r}
param_draws %>%
  filter(param == "delta_at") %>%
  ggplot(aes(x=value))+
  geom_density(alpha = 0.4, linewidth = 0.8, fill = "steelblue", color = "steelblue") +
  labs(x = "Delta AT", y = "density") +
  theme_classic() +
  facet_wrap(~param) +
  xlim(.2,.5)


```

## Inferring Cell Counts

Fitting the model also provides probability distributions over each
unobservable but deterministic quantity. These are quantities that are
downstream of the sampled parameters. For this model, normalized cell
counts $S$ (sensitive) and $R$ (resistant) are unobserved states
determined by the other parameters of the Lotka-Volterra system at each
time point.

This is convenient, because it allows us to infer the cell counts of
sensitive and resistant cells according to how probable they are, given
only the noisy biomarker and the mathematical assumptions. This means
that I get a probability distribution over probable values of sensitive
and resistant cells at each point in time.

In the following plot, I overlay 95% credible intervals of sensitive and
resistant cells at each time point. I've also removed observations and
the inferred total count for legibility. My primary observation is that the
uncertainty in cell counts is not constant throughout the trajectories, but fluctuates quite a bit for the sensitive population and increases for the resistant population.

```{r pred_sr, fig.height=5, fig.cap="*Inferred cell counts of sensitive (S) and resistant (R) at each timepoint. Ribbons around trajectories denote 95% credible intervals, and shaded regions denote treatment was given. Ticks on the x-axis indicate a Prostate Specific Antigen measurement was made.*"}
max_fit <- readRDS(file.path(CACHE_DIR, "p12_t28.rds"))

cell_draws <- spread_draws(max_fit, S_pred[i], R_pred[i]) %>%
  mutate(day   = psa_data$day[i],
         total = S_pred + R_pred)

cell_sum <- cell_draws %>%
  group_by(i, day) %>%
  summarise(
    S_med = median(S_pred), S_lo = quantile(S_pred, 0.025), S_hi = quantile(S_pred, 0.975),
    R_med = median(R_pred), R_lo = quantile(R_pred, 0.025), R_hi = quantile(R_pred, 0.975),
    T_med = median(total),  T_lo = quantile(total,  0.025), T_hi = quantile(total,  0.975),
    .groups = "drop"
  )

p_SR <- ggplot(cell_sum, aes(x = day)) +
  geom_rect(data = trt_abs, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "steelblue", alpha = 0.12, inherit.aes = FALSE) +
  geom_rug(data = psa_data, aes(x = day), inherit.aes = FALSE, sides = "b", colour = "grey40", linewidth = 0.4) +
  geom_ribbon(aes(ymin = S_lo, ymax = S_hi), fill = "steelblue", alpha = 0.2) +
  geom_ribbon(aes(ymin = R_lo, ymax = R_hi), fill = "tomato",    alpha = 0.2) +
  geom_line(aes(y = S_med, colour = "S"),   linewidth = 0.9) +
  geom_line(aes(y = R_med, colour = "R"),   linewidth = 0.9) +
  scale_colour_manual(values = c("S" = "steelblue", "R" = "tomato", "S+R" = "grey30"),
                      name = NULL) +
  labs(x = "Day", y = "Normalized cell count") +
  theme_minimal()

p_SR
```

# Modelling Uncertainty Trough Time

To model how uncertainty changes over time, I fit the same model up to
each $t$. I then
extract posterior estimates for each fit. In other words, there is a
distribution over each parameter for each fit. This allows us to plot
credible intervals for each parameter at each fit and examine how
uncertainty changes with increasing observations.

```{r param_ci}
param_ci_df <- map_dfr(2:N_obs, function(t) {
  cache_path <- file.path(CACHE_DIR, sprintf("p12_t%02d.rds", t))
  fit <- tryCatch(readRDS(cache_path), error = function(e) NULL)
  if (is.null(fit)) return(NULL)
  spread_draws(fit, cost, turnover, n0, rFrac, sigma) %>%
    select(cost, turnover, n0, rFrac, sigma) %>%
    pivot_longer(everything(), names_to = "param", values_to = "value") %>%
    group_by(param) %>%
    summarise(
      med   = median(value),
      lo    = quantile(value, 0.025),
      hi    = quantile(value, 0.975),
      .groups = "drop"
    ) %>%
    mutate(t = t, day = psa_data$day[t])
})

p_param_ci <- ggplot(param_ci_df, aes(x = day)) +
  geom_rect(data = trt_abs, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "steelblue", alpha = 0.12, inherit.aes = FALSE) +
  geom_ribbon(aes(ymin = lo, ymax = hi), fill = "tomato", alpha = 0.25) +
  geom_line(aes(y = med), colour = "tomato", linewidth = 0.9) +
  geom_rug(data = psa_data, aes(x = day), inherit.aes = FALSE, sides = "b", colour = "grey40", linewidth = 0.4) +
  facet_wrap(~param, scales = "free_y", ncol = 2) +
  labs(x = "Day", y = NULL) +
  theme_minimal()

p_param_ci
```

While the uncertainties here also depend on the chosen priors,
which likely need reworking from someone who knows oncology, I find it interesting that only for $\sigma$ and $N_0$ does
the uncertainty appear to decrease in a straightforward "slow but steady
decrease". For *cost*, *turnover* and *initial resistant fraction*, the
uncertainty seems constant until it radically diminishes. I assume the
explanation for this is that individual observations can all of a sudden
rule out a lot of possible parameter combinations.

The following plot shows the trajectory in uncertainty for the Delta AT Score.

```{r extract_delta_at, include=FALSE}
set.seed(1)
prior_draws <- tibble(
  cost     = rbeta(N_PRIOR, 1, 1),
  turnover = rbeta(N_PRIOR, 2, 5),
  n0       = rbeta(N_PRIOR, 1, 1),
  rFrac    = rbeta(N_PRIOR, 1, 30)
) %>%
  mutate(delta_at = calc_delta_at(turnover * rS, k, K,
                                   (1 - cost) * rS, n0 * rFrac))

prior_delta <- prior_draws$delta_at

post_delta_list <- map(1:N_obs, function(t) {
  cache_path <- file.path(CACHE_DIR, sprintf("p12_t%02d.rds", t))
  fit <- tryCatch(readRDS(cache_path), error = function(e) NULL)
  if (is.null(fit)) return(NULL)
  draws <- tryCatch(rstan::extract(fit)$cost, error = function(e) NULL)
  if (is.null(draws) || length(draws) == 0) return(NULL)
  spread_draws(fit, cost, turnover, n0, rFrac) %>%
    mutate(delta_at = calc_delta_at(turnover * rS, k, K,
                                     (1 - cost) * rS, n0 * rFrac)) %>%
    pull(delta_at)
})

delta_df <- map_dfr(1:N_obs, function(t) {
  post <- post_delta_list[[t]]
  if (is.null(post)) return(NULL)
  tibble(t = t, day = psa_data$day[t], delta_at = post[is.finite(post)])
})

kl_df <- map_dfr(1:N_obs, function(t) {
  post_t <- post_delta_list[[t]]
  if (is.null(post_t)) return(NULL)
  kl_prior_post <- kl_div(post_t, prior_delta)
  kl_seq <- if (t > 1 && !is.null(post_delta_list[[t-1]])) {
    kl_div(post_t, post_delta_list[[t-1]])
  } else NA_real_
  tibble(t           = t,
         day         = psa_data$day[t],
         kl_vs_prior = kl_prior_post,
         kl_seq      = kl_seq)
}) %>%
  mutate(days_since_prev = day - lag(day),
         kl_seq_per_day  = kl_seq / days_since_prev)
```

```{r}
ci_df <- map_dfr(1:N_obs, function(t) {
  post <- post_delta_list[[t]]
  if (is.null(post)) return(NULL)
  fin <- post[is.finite(post)]
  if (length(fin) == 0) return(NULL)
  tibble(t   = t,
         day = psa_data$day[t],
         med = median(fin),
         lo  = quantile(fin, 0.025),
         hi  = quantile(fin, 0.975),
         width = quantile(fin, 0.975) - quantile(fin, 0.025))
})

p_ci <- ggplot(ci_df, aes(x = day)) +
  geom_rect(data = trt_abs, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "steelblue", alpha = 0.12, inherit.aes = FALSE) +
  geom_ribbon(aes(ymin = lo, ymax = hi), fill = "tomato", alpha = 0.25) +
  geom_line(aes(y = med), colour = "tomato", linewidth = 0.9) +
  geom_rug(data = psa_data, aes(x = day), inherit.aes = FALSE, sides = "b", colour = "grey40", linewidth = 0.4) +
  labs(x = NULL, y = "delta_AT") +
  theme_minimal()

p_ci
```

Again the uncertainty doesn't appear to decrease in a straightforward
manner. Rather it is extremely large, spanning from 4x to -4x
improvement until around day 300. This is just a case study, i.e. one
patient and one specific mathematical model, but if it is
representative, this makes me think that in order to get an accurate
picture of the Delta AT Score, I think measurements need to be sampled
carefully, and it might be necessary to proactively perturb the
Lotka-Volterra system to make observations that meaningfully
reduce the uncertainty. This could move treatment planning into the
realm of sequential decision-making under uncertainty, where one would
have to balance immediate reward and gaining information about the
system.

## KL-Divergence as a Measure of Information Gain

The above plot shows credible intervals which are easy to interpret: wider equals more uncertainty. But they hide the true shape of the distribution.

Crucially, I find that the distribution of the Delta AT Score
is multimodal for patient 12 given the chosen mathematical model.

```{r plot_delta_densities}
p_dens <- delta_df %>%
  ggplot(aes(x = delta_at, colour = day, fill = day, group = t)) +
  geom_density(alpha = 0.08, linewidth = 0.4) +
  scale_colour_viridis_c(option = "plasma", direction = -1, guide = "none") +
  scale_fill_viridis_c(option = "plasma", direction = -1, name = "day",
                       breaks = range(delta_df$day),
                       labels = c(as.character(min(psa_data$day)), as.character(max(psa_data$day)))) +
  labs(x = "delta_AT", y = "density") +
  xlim(-1, 2) +
  theme_minimal()

p_dens
```

I'd therefore argue that a more principled measure of information gain in the Delta AT Score is the KL divergence between posterior densities at consecutive observations rather than the width of a credible interval. The following plot shows exactly that.
```{r kl_seq, fig.height=4}
kl_df %>%
  filter(!is.na(kl_seq)) %>%
  ggplot(aes(x = day, y = kl_seq)) +
  geom_col(fill = "grey") +
  labs(x = "Day of observation", y = "Information Gain (nats)") +
  theme_bw()
```

From this it is also clear that information gain in the Delta AT Score does not
follow a straightforward pattern where more observations generally lead
to more information. Instead, some observations radically change the
posterior density for the Delta AT Score. The uneven spacing of observations does
complicate hypothesizing about what might be driving information gain.
My take is that the model can't decipher enough about cost nor the initial resistant population until resistant cells make up a substantial part of the PSA signal.

```{r}
p_SR + geom_col(data=kl_df, aes(x = day, y = kl_seq), alpha = .3)
```

Assuming that this case study in patient 12 is representative, I think
targeting information gain could be important in planning treatments
given the circumstances: a relatively complicated mathematical model,
only one presumably noisy output (PSA), and few observations. However,
our ability to perturb the system by applying treatment might allow us
to gain an accurate picture of the system. I suspect two things are
going on when information gain peaks at the beginning of the second
treatment block: i) the model is only able to differentiate how much of
the signal comes from resistant cells when treatment is being applied,
since they react differently to treatment than sensitive cells; ii) the
resistant cells have been freed from the competitive pressure of the
sensitive cells during the second treatment block, their numbers
therefore grow, and the model substantially updates the *cost*
parameter. This update then cascades through to large information gain
about the Delta AT Score.

# All in All

Adaptive therapy is a cancer therapy that attempts to manipulate
evolutionary dynamics to stabilize rather than eradicate cancers.
@gallagher propose a metric to determine the potential benefit of
adaptive therapy over standard care for prostate cancer patients. They
explicitly state that they leave the quantification of uncertainty in
these measures for future work. In this notebook, I have used Bayesian
inference to quantify the uncertainty in one of those metrics: the Delta AT Score.
I have further conducted a case study tracing the *trajectory of
uncertainty* for each added observation for a single prostate cancer
patient. I argue that uncertainty in this measure does not decrease
straightforwardly with each additional observation. Instead, it may be necessary to proactively search out observations
that reduce the uncertainty, and lessons from sequential decision-making
under uncertainty may apply.

## Suggested Next Steps

I would suggest that someone decides on a provisional candidate
mathematical model of prostate cancer dynamics. Then one could simulate
noisy emissions from this system. For this to be possible, one would
have to map cell counts to Prostate Specific Antigen measurements. I
assume there is substantial noise in this mapping, but again I'm
definitely not an expert on this. My first instinct would be to do
Bayesian model comparison across different likelihoods on the Bruchovsky
data. One could then place rewards on extending time-to-progression and
on making observations that lead to better information about the system.
KL divergence strikes me as an obvious starting choice for measuring
information gain. Optimal strategies under this scenario would likely
require flexible observation scheduling, so that measurements are made
when high information gain is predicted.

For another project, I formulated the treatment problem as a discrete state-space and compared two canonical Partially Observable Markov Decision Process
solvers on their ability to extend time to progression. I would be
interested in reformulating this as a continuous state-space problem if
anyone cares to fund and collaborate. I don't know if the problem is
tractable for a continuous state-space, but @saporta2024 appear to have
found solutions to a somewhat similar problem.

# Rerunning this Notebook
Given the hardware used, you might have to change the number of cores used for running stan. See the fitting script. Fitting the entire procedure took a while (approx. an hour) on a M1 MacBook Air for reference. You will need to have the following in place:
- Functioning version of R and relevant packages installed
- Stan properly setup
- The [patient data](https://www.nicholasbruchovsky.com/clinicalResearch.html).

# References


